Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue

### 2020-12-17 reports/rmarkdown/03_integrated_datasets.Rmd
adds figure1 in manuscript

samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
metadata <- metadata[ samples_to_use , ]
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)

#Compute sample-wise correlations
cors <- cor(all_microbiome, method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

@@

#Graph construction  
k <- 10 # k for KNN
P <- 0.5# prunning threshold of SNN

@@

k=20
P=0.1
snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

@@

@@

g <- graph_from_adjacency_matrix(snnm,mode = "undirected",diag = F,weighted = T)

mypar(2,2,mar=c(0,0,2,10))

set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)
metadata$joint_clustering <- cl$membership

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)

@@

clean_Ftest <- Ftest
# clean_Ftest <- ( clean_Ftest + t(clean_Ftest) ) / 2
clean_Ftest <- -log10(clean_Ftest)
clean_Ftest[clean_Ftest > 5] <- 5
# clean_Ftest[ clean_Ftest > 0.05 ] <- 1

@@

#Graph construction  
k <- 5 # k for KNN
P <- 0.1 # prunning threshold of SNN

knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

# k=10
# P=0.1
snn <- dbscan::sNN(x = as.dist( (1-cors)/2 ), k = k, jp = FALSE, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
snnm[ snnj <= ceiling(k*P) ] <- 0

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
```

# Visualise the data

```{r, echo = FALSE,fig.width=10,fig.height=9}
mypar(2,2,mar=c(0,0,2,10))

set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=1)

# cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)
# 
table(cl$membership)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)
```

@@

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}

@@

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}
```

>>>>

samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
metadata <- metadata[ samples_to_use , ]
saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
# all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
mypar(mar=c(2,15,2,2))
top_var <- apply(all_microbiome,1,var)
top_var_t <- sort(top_var,decreasing = T)[100]
is_top_var <- top_var > top_var_t 
barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
#Compute sample-wise correlations
cors <- cor(all_microbiome[is_top_var,], method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

@@

#Graph construction  
k <- 15 # k for KNN
P <- 0.1# prunning threshold of SNN

@@

# k=10
# P=0.1
snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

@@

N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < 0.2] <- 0
colnames(J) <- colnames(all_microbiome)
rownames(J) <- colnames(all_microbiome)
# image(J[,nrow(nn):1],col = colorRampPalette(c("grey95","black"))(90),axes=F )
# set.seed(1)
# g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
# plot( g , vertex.label.cex=0.000001 , vertex.color = "grey" , 
#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
#       vertex.size=10 )

@@

g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)

mypar(3,3,mar=c(0,0,2,0))

set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)
metadata$joint_clustering <- cl$membership
set.seed(1)
U <- uwot::umap(adj,n_neighbors = 30)
h <- hclust(as.dist(adj),method = "ward.D2")
# metadata$joint_clustering <- cutree(h,k = 5)
write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding\nlouvain",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups",
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
       legend = levels(factor(metadata$group_M4_CVLv3)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

@@

clean_Ftest <- Ftest
# clean_Ftest <- ( clean_Ftest + t(clean_Ftest) ) / 2
clean_Ftest <- -log10(clean_Ftest)
clean_Ftest[clean_Ftest > 5.5] <- 5.5
# clean_Ftest[ clean_Ftest > 0.05 ] <- 1

write.csv2(clean_Ftest , "../../results/metadata_association_results.csv",row.names = T)

@@

#Graph construction 
k_nearest_neighbours <- 10 # k for KNN
snn_threshold <- 0.1 # prunning threshold of SNN

#Compute kNN graph
knn <- RANN::nn2( (1-cors)/2 ,k = k_nearest_neighbours)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Calculate SNN and prune based on Jaccard index
N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < snn_threshold] <- 0
rownames(J) <- rownames(cors)
colnames(J) <- colnames(cors)

#Build a graph object, perform clustering and define the force-directed layout
g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,"../../results/bacteria_SNN_graph.csv",row.names = T)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=1)

#Define layout with UMAP
mU <- uwot::umap((1-cors)/2, n_neighbors = 10)

#Plot
mypar(3,3,mar=c(0,0,2,0))

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
    vertex.size=10, main="graph layout FR",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
    vertex.size=10, main="UMAP",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
par(mar=c(0,0,0,0))
empty_plot()
legend("topleft",title = "bact. communities",
     legend = levels(factor(cl$membership)),
     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)

@@

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),col = pal,
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}

@@

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),col=pal,
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}
```

### 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds figure1 in manuscript

---
title: "MANUSCRIPT TITLE"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
# library(ggthemes)
library(dplyr)
library(stringr)
remotes::install_github("czarnewski/niceRplots")
library(niceRplots)
```

@@

```{r, echo = FALSE, eval = TRUE}
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
patient_groups <- metadata$joint_clustering
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
```

**Joint analysis of vaginal microbiome reveals XXXXXXXXXXXXXX**

@@

@@

Bumpus focused his study on the House Sparrow (*Passer domesticus*; see Figure 1), which has a very wide global distribution. It is native to Europe and Asia, but not the Americas where Bumpus collected his original study [@Bumpus1898]. In addition to measuring total length and survival for 136 sparrows, Bumpus measured sparrow sex, wingspan, and mass, and also the length of each sparrow's head, humerus, tibiotarsus, skull, and sternum. While modern ornithologists believe that the total body length measurement that I will use today is subject to high observational error [@Johnston1972], it will be more than sufficient for demonstrating Rmarkdown.

<!--- Note that the image below is in the public domain --->
<!--- https://commons.wikimedia.org/wiki/File:Lto-tpbo-passer-domesticus-05.JPG --->
<center>

![*Passer domesticus*](images/sparrow.jpg)

</center>

I performed an independent two-sample student's t-test on sparrow total body length to test whether or not sparrows that died in the 1898 storm were larger than sparrows that survived. I assume that both groups of sparrows (dead and living) have equal variances, so the test statistic $t$ is calculated as follows,

$$t = \frac{\bar{X}_{1} - \bar{X}_{2}} {s_{p} \times \sqrt{\frac{1}{n_{1}} + \frac{1}{n_{2}}}}.$$

In the above, $\bar{X}_{1}$ and $\bar{X}_{2}$ are the mean of the samples of sparrows that died and lived, respectively. Similarly, $n_{1}$ and $n_{2}$ are the sample sizes of sparrows that died and lived, and $s_{p}$ is the pooled sample mean, which is calculated as follows,

$$s_{p} = \sqrt{\frac{s^{2}_{X_{1}} + s^{2}_{X_{2}}}{2}}.$$

In the above, the $s^{2}_{X_{1}}$ and $s^{2}_{X_{2}}$ are the sample standard deviations for sparrows that died and lived, respectively. I conduceted the two sample t-test using the `t.test` function in R.

@@

>>>>

---
title: "Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(out.width = '100%',
                      dpi=600,
                      fig.width=10,
                      echo = FALSE,
                      # fig.pos = 'p', # Places figures on their own pages
                      eval = TRUE, 
                      results = 'hide',
                      warning = FALSE,
                      message = FALSE )

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
library(igraph)
library(dplyr)
library(stringr)
library(rafalib)
remotes::install_github("czarnewski/niceRplots")
library(niceRplots)

#color palettes
pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

@@

```{r, echo = FALSE, eval = TRUE}
# Load reasult from the analysis
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
patient_groups <- metadata$group_M4_CVLv3
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
clean_Ftest <- read.csv2("../../../results/metadata_association_results.csv",row.names = 1)
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")

```

**Joint analysis of vaginal microbiome reveals distinct patient subgroups**

@@

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

**Indentification of bacterial communities metabolic processes linked to Lactobacilli**

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

**Indentification of bacterial communities metabolic processes linked to Lactobacilli**

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

@@

@@

FIGURES (MAIN)
================================================================================

## Figure 1

```{r, echo = FALSE, eval = TRUE , fig.asp=6/10}
# Define Layout and set labels
layout(matrix(c(1,1,1,1,      3,3,3,3,          4,4,4,4,
                2,2,2,2,      3,3,3,3,          5,5,5,5,
                6,6,6,6,      9,9,9,9,          9,9,9,9,
                7,7,7,7,     10,10,10,10,     10,10,10,10,
                8,8,8,8,     10,10,10,10,     10,10,10,10 ) , ncol= 12,byrow = T),
       widths = 1,heights = c(1,1,0.5,0.5,0.5))
# layout.show(29)
figlabels <- letters

#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0,0,0,0))
fig1a <- magick::image_read("schematics.pdf",density = 900)
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

####################
# PATIENT GROUPING #
####################
g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

par(mar=c(0,0,0,0))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$joint_clustering)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10,main="", 
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend("topright",title.adj = 0,
       legend = levels(factor(metadata$joint_clustering)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

##############################
# METADATA CROSS-ASSOCIATION #
##############################
par(mar=c(12,1,1,12))
image( t(clean_Ftest [nrow(clean_Ftest):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(clean_Ftest)) ,
       labels = rownames(clean_Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(clean_Ftest)) , par("usr")[3]-.02 ,
      labels = rownames(clean_Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(0,0,0,0))
empty_plot(frame = T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

empty_plot(frame = T)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Monbilicus",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardanerella",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,1,1,1))

shann <- vegan::diversity(t(x),index = "shannon")
shann <- t(as.matrix(shann))
rownames(shann) <- "Shannon"
barlist( data = shann,
         main = "Shannon",
         labels = "",
         genes = c("Shannon"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(4,4,1,1))
temp <- t(t(2^x-1)/colSums(2^x-1))
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
o <- order( rowSums(temp) , decreasing = F)
sample_o <- order( as.numeric(metadata$group_M4_CVLv3)+(shann/max(shann)/2) )
temp <- temp[o,sample_o]*100
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i", col = pal[factor(rownames(temp))] )

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]
```

**Figure 1. Identification of patient groups.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

## Figure 2

```{r, echo = FALSE, eval = TRUE}
# Define Layout and set labels
layout(matrix(c(1) , ncol= 1,byrow = T),
       heights = c(1))
figlabels <- letters

```

**Figure 2. Identification and characterization of vaginal bacterial communities.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

## Figure 3

```{r, echo = FALSE, eval = TRUE}
# Define Layout and set labels
layout(matrix(c(1) , ncol= 1,byrow = T),
       heights = c(1))
figlabels <- letters

```

**Figure 1. Identification of patient groups.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

FIGURES (SUPPL)
================================================================================

## Figure S1

## Figure S2

## Figure S3

## Figure S4

TABLES (MAIN)
================================================================================

## Table 1

## Table 2

## Table 3

TABLES (SUPPL)
================================================================================

## Table S1

## Table S2

## Table S3
